### 2025-01-13 report/report_template.qmd
folders

---
title: "Group assignment: Is High-Dimensional Analysis Worth it?"
format: 
  pdf:
    toc: false
    number-sections: true
    documentclass: article
    geometry: "margin=4cm"
    bibliography: bibliography.bib
abstract: |
    You can write the abstract at the end. Do it approximately like this and you have a nice flow: one sentence about data context. One sentence about research problem. One sentence about research question. One sentence about methods compared. One sentence about results. One sentence about conclusion.
author: 
  - Carlos Perengano
  - Erika Mustermann
  - Jan Kowalski
---

# Introduction

::: {.callout-note}
Max 700 words
:::

Provide a description of your data and relevant context on the research problem it might have been designed to solve in 2 to 3 paragraphs. Maybe referring to a figure in a manner such as @fig-introimage.

```{r}
#| label: fig-introimage
#| fig-cap: A boxplot comparing the mpg of cars for different numbers of cylinders.
#| out-width: "80%"
#| echo: false
boxplot(mpg ~ cyl, data = mtcars)
```

Write max 2 paragraphs describing the research problem and leading to the methodological research question. The question should revolve around comparing high-dimensional vs not high-dimensional methods.

# Methods

::: {.callout-note}
500-600 words
:::

Description of methods you use in maximum 3 paragraphs. You might want to refer to literature which you can find in the file `bibliography.bib` e.g., @hastie2015statistical or @james2013introduction. Use bibtex notation in the `.bib` file.

Also describe your comparison or validation strategy in 1 or 2 paragraphs. Answer the question: how are you going to compare your chosen methods against each other? How are you going to conclude which is better for this dataset / research purpose?

# Results

::: {.callout-note}
Max 500 words
:::

Here, provide a description of __relevant__ results in 2 to 3 paragraphs. You are invited to use images / plots to help communicate your results, but keep this section legible and stick to the main points! In your results section, you should refrain from giving too much discussion / explanation; be quite factual.

# Conclusion

::: {.callout-note}
Max 700 words
:::

Start with a short summary of what you have done in 1 paragraph ("in this report, we have...").

Then, write your conclusion in which you answer your research question in 1 paragraph ("For this dataset, we have found that the high-dimensional analysis improved / did not improve over ..." )

Then, contextualize your conclusion by providing some discussion additional interpretation in approx. 3 paragraphs. Feel free to refer to additional literature if you like. Suggestions for future research might be nice to add.

# References
<!-- Don't put anything here, your references will be inserted automatically at the end -->

### 2025-01-13 code/analysis.qmd
add reference model

---
title: "INFOMDA2: Group Assignment"
author: "Martijn van Dam, Jonathan Koop, Merlin Urbanski"
format: html
---

# 1. Preparations

## 1.1 Load Packages

```{r}
library(pacman)

p_load("readr", "tidyverse", "caret")
```

## 1.2 Load Data

```{r}
data <- read.csv("../raw_data/data.csv")
```

# 2. Data Exploration

First, we can explore what the data looks like by looking at the first few rows of the data set.

```{r}
head(data)
```

We can also look at the structure of the data set to see what kind of variables we are dealing with.

```{r}
str(data)
```

As can be seen from the output above, all variables except for `ID` are either numeric or integers. Since we have a categorical outcome, we have to transform it to a factor

```{r}
data <- data %>%
  mutate(class = as.factor(class)) # change class to factor
```

# 3. Benchmark Approach (Not High-Dimensional)

## 3.1 Variable/Feature Selection

In our benchmark approach, which does not properly deal with the high dimensionality of our data, we use $t$-tests for differences in group means in the dependent variable `class` for each of the independent variables.

First, we check which variables lead to a $p$-value below 0.05.

```{r}
sig_05 <- data %>%
  select(-ID) %>% # remove ID column
  summarise(across(-class, ~ t.test(.x ~ class)$p.value)) %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("variable") %>%
  filter(V1 < 0.05)
```

The code above performs t-tests for each variable in the data set and filters out the variables for which the $p$-value is below 0.05. When we run the code, we see that this approach does not effectively reduces the number of variables, as we still have `r nrow(simple_data)` variables left. This leads to an unidentifiable model, because we still have more variables than observations. For that reason, we reduce the threshold to 0.01.

```{r}
sig_01 <- data %>%
  select(-ID) %>% # remove ID column
  summarise(across(-class, ~ t.test(.x ~ class)$p.value)) %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("variable") %>%
  filter(V1 < 0.01)
```

We still have too many significant variables. Consequently, we reduce the threshold further.

```{r}
sig_001 <- data %>%
  select(-ID) %>% # remove ID column
  summarise(across(-class, ~ t.test(.x ~ class)$p.value)) %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("variable") %>%
  filter(V1 < 0.001)
```

As we can see, using $p<0.001$ as a threshold reduces the number of variables to `r nrow(sig_001)`. As this is less than the number of observations, we can now proceed with the model building.

## 3.2 Model Building

### 3.2.1 Subset Data

```{r}
data_benchmark <- data %>%
  select(sig_001$variable, class) # select only significant variables and DV
```

### 3.2.2 Logistic Regression with Cross-Validation

```{r}
set.seed(123)
model_benchmark <- train(class ~ ., data = data_benchmark, method = "glm", family = "binomial", trControl = trainControl(method = "cv", number = 10))

summary(model_benchmark)
```

### 2025-01-13 code/analysis.qmd
add high-dimensional model

@@
 ---
+
+```{r}
+#| echo: false
+options(scipen = 999)
+```
 
@@
 
-p_load("readr", "tidyverse", "caret")
+p_load("readr", "tidyverse", "caret", "kableExtra")
 ```
@@
   mutate(class = as.factor(class)) # change class to factor
+
+data$class <- as.factor(data$class)
 ```
@@
 The code above performs t-tests for each variable in the data set and filters out the variables for which the $p$-value is below 0.05.
-When we run the code, we see that this approach does not effectively reduces the number of variables, as we still have `r nrow(simple_data)` variables left.
+When we run the code, we see that this approach does not effectively reduces the number of variables, as we still have `r nrow(sig_05)` variables left.
 This leads to an unidentifiable model, because we still have more variables than observations.
@@
 summary(model_benchmark)
+model_benchmark$results
 ```
+
+Even though we have reduced the number of variables to `r nrow(sig_001)`, the algorithm still does not seem to converge. In the li
+
+## 3.3 7 Most Significant Variables
+
+```{r}
+sig_7 <- sig_001 %>%
+  top_n(7, V1) %>%
+  pull(variable)
+
+data_7 <- data %>%
+  select(sig_7, class) %>% 
+  train(class ~ ., data = ., method = "glm", family = "binomial", trControl = trainControl(method = "cv", number = 10))
+
+data_7$results
+```
+
+# 4. Principal Component Regression (High-Dimensional Approach)
+
+## 4.1 Principal Component Analysis
+
+### 4.1.1 Identify Principal Components
+
+```{r}
+pca <- data %>%
+  select(-ID, -class) %>%
+  prcomp(center = TRUE, scale. = TRUE)
+```
+
+### 4.1.2 Scree Plot
+
+```{r}
+screeplot(pca, type = "lines", main = "Scree Plot")
+```
+
+Analyzing the scree plot, we can see that the first component explains most of the variance in the data, explaining three times as much variance as the second component.
+The third component explains even less variance.
+This suggests that we can reduce the dimensionality of the data by only using the first component, as the elbow is already at the second component.
+
+### 4.1.3 Variances Explained
+
+Only seeing the absolute eigenvalue of the components does not give us a good idea of how much variance each component explains.
+Therefore, we calculate the proportion of variance explained by each component and the cumulative proportion of variance explained.
+
+```{r}
+pca_table <- tibble(
+  component = seq_along(pca$sdev),       # Component numbers
+  eigenvalue = pca$sdev^2                 # Variances (eigenvalues)
+) %>%
+  mutate(
+    proportion = eigenvalue / sum(eigenvalue),  # Proportion of variance explained
+    cumulative = cumsum(proportion)        # Cumulative variance explained
+  ) %>%
+  round(., 3)
+```
+
+As can be seen in the table below, the first component explains only `r pca_table$proportion[1]` of the variance.
+This implies that simply using the first component would not be a good idea, as it does not explain enough variance.
+We can see that even the first 10 components only explain `r pca_table$cumulative[10]` of the variance.
+This suggests that we should use more than just the first components.
+Therefore, the eigenvalue greater than 1 rule may perform better in this case.
+
+```{r}
+pca_table %>%
+  filter(eigenvalue > 1) %>%
+  kable("html") %>%
+  kable_styling("striped", full_width = F)
+```
+
+### 4.1.4 Eigenvalue Greater Than 1 Rule
+
+```{r}
+pc_greaterthan1 <- pca_table %>%
+  filter(eigenvalue > 1) %>%
+  pull(component) %>% 
+  paste0("pc", .)
+```
+
+## 4.2 Model Building
+
+### 4.2.1 Subset Data
+
+```{r}
+data_pca <- bind_cols(
+  class = data$class,  # Add class column
+  as_tibble(pca$x)     # Convert PCA matrix to tibble
+) %>%
+  set_names(c("class", paste0("pc", 1:ncol(pca$x)))) 
+```
+
+### 4.2.2 Scree Plot
+
+```{r}
+logit_scree <- data_pca %>%
+  select(pc1, class) %>%
+  train(
+    class ~ .,
+    data = .,
+    method = "glm",
+    family = "binomial",
+    trControl = trainControl(method = "cv", number = 10)
+  )
+
+summary(logit_scree)
+logit_scree$results
+```
+
+### 4.2.3 Eigenvalue Greater Than 1 Rule
+
+```{r}
+logit_greaterthan1 <- data_pca %>% 
+  select(all_of(pc_greaterthan1), class) %>%
+  train(
+    class ~ .,
+    data = .,
+    method = "glm",
+    family = "binomial",
+    trControl = trainControl(method = "cv", number = 10)
+  )
+
+summary(logit_greaterthan1)
+logit_greaterthan1$results
+```
+
+```{r}
+logit_7 <- data_pca %>% 
+  select(pc1, pc2, pc3, pc4, pc5, pc6, pc7, class) %>%
+  train(
+    class ~ .,
+    data = .,
+    method = "glm",
+    family = "binomial",
+    trControl = trainControl(method = "cv", number = 10)
+  )
+
+summary(logit_7)
+logit_7$results
+```

### 2025-01-15 code/analysis.qmd
update analysis

@@
 
-Even though we have reduced the number of variables to `r nrow(sig_001)`, the algorithm still does not seem to converge. In the li
+Even though we have reduced the number of variables to `r nrow(sig_001)`, the algorithm still does not seem to converge. The issue here
 
@@
 
+### Iteratively Add Components
+
 ```{r}
-logit_7 <- data_pca %>% 
-  select(pc1, pc2, pc3, pc4, pc5, pc6, pc7, class) %>%
+logit_results <- rep(NA, length(pc_greaterthan1))
+
+for (i in 1:length(pc_greaterthan1)) {
+  logit <- data_pca %>% 
+    select(all_of(pc_greaterthan1[1:i]), class) %>%
     train(
@@
       family = "binomial",
-    trControl = trainControl(method = "cv", number = 10)
+      trControl = trainControl(method = "LOOCV")
     )
 
-summary(logit_7)
-logit_7$results
+  logit_results[i] <- logit$results$Accuracy
+
+}
+
+ggplot(data = data.frame(x = 1:length(pc_greaterthan1), y = logit_results), aes(x = x, y = y)) +
+  geom_line() +
+  labs(x = "Number of Components", y = "Accuracy", title = "Accuracy by Number of Components") +
+  theme_minimal()
 ```
+
+Only keep components that increase accuracy
+
+```{r}
+diff <- c(0, diff(logit_results))
+
+index <- which(diff >= 0)
+```
+
+```{r}
+logit_results_incr <- rep(NA, length(index))
+
+for (i in seq_along(index)) {
+  logit <- data_pca %>% 
+    select(all_of(pc_greaterthan1[index[1:i]]), class) %>%
+    train(
+      class ~ .,
+      data = .,
+      method = "glm",
+      family = "binomial",
+      trControl = trainControl(method = "LOOCV")
+    )
+
+  logit_results_incr[i] <- logit$results$Accuracy
+
+}
+
+diff_incr <- c(0, diff(logit_results_incr))
+
+index_incr <- index[which(diff_incr >= 0)]
+
+logit_results_incr2 <- rep(NA, length(index_incr))
+
+for (i in seq_along(index_incr)) {
+  logit <- data_pca %>% 
+    select(all_of(pc_greaterthan1[index_incr[1:i]]), class) %>%
+    train(
+      class ~ .,
+      data = .,
+      method = "glm",
+      family = "binomial",
+      trControl = trainControl(method = "LOOCV")
+    )
+
+  logit_results_incr2[i] <- logit$results$Accuracy
+
+}
+
+diff_incr2 <- c(0, diff(logit_results_incr2))
+
+index_incr2 <- index_incr[which(diff_incr2 >= 0)]
+
+logit_results_incr3 <- rep(NA, length(index_incr2))
+
+for (i in seq_along(index_incr2)) {
+  logit <- data_pca %>% 
+    select(all_of(pc_greaterthan1[index_incr[1:i]]), class) %>%
+    train(
+      class ~ .,
+      data = .,
+      method = "glm",
+      family = "binomial",
+      trControl = trainControl(method = "LOOCV")
+    )
+
+  logit_results_incr3[i] <- logit$results$Accuracy
+
+}
+```
+
+```{r}
+iterations <- 10
+
+# Initialize index_incr for the first iteration
+index_incr <- index
+
+# Store results of each iteration
+logit_results_all <- list()
+diff_results_all <- list()
+
+for (iteration in 1:iterations) {
+
+  logit_results <- rep(NA, length(index_incr))
+
+  for (i in seq_along(index_incr)) {
+    logit <- data_pca %>% 
+      select(all_of(pc_greaterthan1[index_incr[1:i]]), class) %>%
+      train(
+        class ~ .,
+        data = .,
+        method = "glm",
+        family = "binomial",
+        trControl = trainControl(method = "LOOCV")
+      )
+
+    logit_results[i] <- logit$results$Accuracy
+  }
+
+  diff_results <- c(0, diff(logit_results))
+
+  # Save results of the current iteration
+  logit_results_all[[iteration]] <- logit_results
+  diff_results_all[[iteration]] <- diff_results
+
+  # Update index_incr for the next iteration
+  index_incr <- index_incr[which(diff_results >= 0)]
+
+  # Break the loop if no more indices are selected
+  if (length(index_incr) == 0) break
+}
+
+# Final result stored in logit_results_all and diff_results_all for all iterations
+
+```
+
+```{r}
+ggplot(data = data.frame(x = 1:length(index_incr), y = logit_results_all[[10]]), aes(x = x, y = y)) +
+  geom_line() +
+  labs(x = "Number of Components", y = "Accuracy", title = "Accuracy by Number of Components") +
+  theme_minimal()
+```
+
+```{r}
+logit <- data_pca %>% 
+      select(all_of(pc_greaterthan1[index_incr]), class) %>%
+      train(
+        class ~ .,
+        data = .,
+        method = "glm",
+        family = "binomial",
+        trControl = trainControl(method = "LOOCV")
+      )
+
+summary(logit)
+```

### 2025-01-15 code/analysis.qmd
change to loocv

@@
 set.seed(123)
-model_benchmark <- train(class ~ ., data = data_benchmark, method = "glm", family = "binomial", trControl = trainControl(method = "cv", number = 10))
+model_benchmark <- train(class ~ ., data = data_benchmark, method = "glm", family = "binomial", trControl = trainControl(method = "LOOCV"))
 
@@
   select(sig_7, class) %>% 
-  train(class ~ ., data = ., method = "glm", family = "binomial", trControl = trainControl(method = "cv", number = 10))
+  train(class ~ ., data = ., method = "glm", family = "binomial", trControl = trainControl(method = "LOOCV"))
 
@@
     family = "binomial",
-    trControl = trainControl(method = "cv", number = 10)
+    trControl = trainControl(method = "LOOCV")
   )
@@
 
-summary(logit)
 ```

### 2025-01-16 code/analysis.qmd
final changes to code, add plots

--- 2025-01-13 code/analysis.qmd
@@
 format: html
+knitr:
+  opts_chunk: 
+    warning: false
 ---
@@
 ```{r}
+# if required, install pacman package
+if (!requireNamespace("pacman", quietly = TRUE)) {
+  install.packages("pacman")
+}
+
 library(pacman)
 
-p_load("readr", "tidyverse", "caret", "kableExtra")
+# load all packages with pacman
+p_load("readr", "tidyverse", "caret", "kableExtra", "pROC")
 ```
@@
 ```{r}
-data <- read.csv("../raw_data/data.csv")
+data <- read.csv("../raw_data/data.csv") # load data
 ```
@@
 ```{r}
-head(data)
+head(data) # show first few rows of data
 ```
@@
 ```{r}
-str(data)
+str(data) # show structure of data
 ```
@@
 ```{r}
-data <- data %>%
-  mutate(class = as.factor(class)) # change class to factor
-
+# Convert class to factor
 data$class <- as.factor(data$class)
@@
   select(-ID) %>% # remove ID column
-  summarise(across(-class, ~ t.test(.x ~ class)$p.value)) %>%
-  t() %>%
+  summarise(across(-class, ~ t.test(.x ~ class)$p.value)) %>% # perform t-tests
+  t() %>% # transpose
   as.data.frame() %>%
-  rownames_to_column("variable") %>%
-  filter(V1 < 0.05)
+  rownames_to_column("variable") %>% # add variable names as a column
+  filter(V1 < 0.05) # filter out variables with p-value < 0.05
 ```
@@
   rownames_to_column("variable") %>%
-  filter(V1 < 0.01)
+  filter(V1 < 0.01) # filter out variables with p-value < 0.01
 ```
@@
   rownames_to_column("variable") %>%
-  filter(V1 < 0.001)
+  filter(V1 < 0.001) # filter out variables with p-value < 0.001
 ```
@@
 
+#### 3.2.2.1 Training the Model
+
 ```{r}
-set.seed(123)
-model_benchmark <- train(class ~ ., data = data_benchmark, method = "glm", family = "binomial", trControl = trainControl(method = "cv", number = 10))
+# | warning: false
 
-summary(model_benchmark)
-model_benchmark$results
+set.seed(123) # set seed for reproducibility
+
+# Train logistic regression model
+model_benchmark <- train(
+  class ~ ., # all significant variables
+  data = data_benchmark,
+  method = "glm",
+  family = "binomial",
+  trControl = trainControl(method = "LOOCV", savePredictions = "final", classProbs = TRUE)) # leave-one-out cross-validation
+
+summary(model_benchmark) # show summary of model
 ```
 
-Even though we have reduced the number of variables to `r nrow(sig_001)`, the algorithm still does not seem to converge. In the li
+Even though we have reduced the number of variables to `r nrow(sig_001)`, the algorithm still does not seem to converge.
 
-## 3.3 7 Most Significant Variables
+#### 3.2.2.2 Making Predictions and Evaluating the Model
 
 ```{r}
-sig_7 <- sig_001 %>%
-  top_n(7, V1) %>%
-  pull(variable)
+# Accuracy and Kappa
+model_benchmark$results
 
-data_7 <- data %>%
-  select(sig_7, class) %>% 
-  train(class ~ ., data = ., method = "glm", family = "binomial", trControl = trainControl(method = "cv", number = 10))
-
-data_7$results
+# ROC Curve and AUC
+predictions_model_benchmark <- model_benchmark$pred
+roc_benchmark <- roc(response = predictions_model_benchmark$obs, predictor = predictions_model_benchmark$P)
 ```
@@
 pca <- data %>%
-  select(-ID, -class) %>%
-  prcomp(center = TRUE, scale. = TRUE)
+  select(-ID, -class) %>% # remove ID and class columns
+  prcomp(center = TRUE, scale. = TRUE) # perform PCA
 ```
@@
 ```{r}
-screeplot(pca, type = "lines", main = "Scree Plot")
+screeplot(pca, type = "lines", main = "Scree Plot") # plot scree plot
 ```
@@
 pca_table %>%
-  filter(eigenvalue > 1) %>%
-  kable("html") %>%
-  kable_styling("striped", full_width = F)
+  filter(eigenvalue > 1) %>% # filter out components with eigenvalue > 1
+  kable("html") %>% # create HTML table
+  kable_styling("striped", full_width = F) # add styling
 ```
@@
 pc_greaterthan1 <- pca_table %>% 
-  filter(eigenvalue > 1) %>%
-  pull(component) %>% 
-  paste0("pc", .)
+  filter(eigenvalue > 1) %>% # filter out components with eigenvalue > 1
+  pull(component) %>% # extract component numbers
+  paste0("pc", .) # create column names
 ```
@@
 ) %>%
-  set_names(c("class", paste0("pc", 1:ncol(pca$x)))) 
+  set_names(c("class", paste0("pc", 1:ncol(pca$x)))) # Rename columns
 ```
 
-### 4.2.2 Scree Plot
+### 4.2.2 Logistic Regression Using Elbow Criterion
+
+#### 4.2.2.1 Training the Model
 
 ```{r}
-logit_scree <- data_pca %>%
+model_scree <- data_pca %>%
   select(pc1, class) %>%
@@
     family = "binomial",
-    trControl = trainControl(method = "cv", number = 10)
+    trControl = trainControl(method = "LOOCV", savePredictions = "final", classProbs = TRUE) # leave-one-out cross-validation
   )
 
-summary(logit_scree)
-logit_scree$results
+summary(model_scree) # show summary of model
 ```
 
-### 4.2.3 Eigenvalue Greater Than 1 Rule
+#### 4.2.2.2 Making Predictions and Evaluating the Model
 
 ```{r}
-logit_greaterthan1 <- data_pca %>% 
+# Accuracy and Kappa
+model_scree$results
+
+# ROC Curve and AUC
+predictions_model_scree <- model_scree$pred
+roc_scree<- roc(response = predictions_model_scree$obs, predictor = predictions_model_scree$P)
+```
+
+### 4.2.3 Logistic Regression Using Eigenvalue Greater Than 1 Rule
+
+#### 4.2.3.1 Training the Model
+
+```{r}
+model_greaterthan1 <- data_pca %>% 
   select(all_of(pc_greaterthan1), class) %>%
@@
     family = "binomial",
-    trControl = trainControl(method = "cv", number = 10)
+    trControl = trainControl(method = "LOOCV", savePredictions = "final", classProbs = TRUE)
   )
 
-summary(logit_greaterthan1)
-logit_greaterthan1$results
+summary(model_greaterthan1)
 ```
 
+#### 4.2.3.2 Making Predictions and Evaluating the Model
+
 ```{r}
-logit_7 <- data_pca %>% 
-  select(pc1, pc2, pc3, pc4, pc5, pc6, pc7, class) %>%
+# Accuracy and Kappa
+model_greaterthan1$results
+
+# ROC Curve and AUC
+predictions_model_greaterthan1 <- model_greaterthan1$pred
+roc_greaterthan1 <- roc(response = predictions_model_greaterthan1$obs, predictor = predictions_model_greaterthan1$P)
+```
+
+# 5. Plots
+
+## 5.1 Scree Plot
+
+```{r}
+# Slice out the first 90 components
+pca_table_90 <- pca_table %>% 
+  slice(1:90)
+
+ggplot(pca_table_90, aes(x = component)) +
+  # Bar plot for variance explained
+  geom_col(aes(y = eigenvalue), fill = "steelblue") +
+
+  # Line + points for cumulative explained variance, scaled to max(eigenvalue)
+  geom_line(aes(y = cumulative * max(eigenvalue)), color = "red", size = 1) +
+  geom_point(aes(y = cumulative * max(eigenvalue)), color = "red", size = 2) +
+
+  # Set up x-axis breaks (optional, e.g. every 5 components)
+  scale_x_continuous(breaks = seq(0, 90, by = 5)) +
+
+  # Primary y-axis is variance; secondary y-axis is the cumulative proportion
+  scale_y_continuous(
+    name = "Variance (Eigenvalue)",
+    sec.axis = sec_axis(
+      trans = ~ . / max(pca_table_90$eigenvalue),
+      name = "Cumulative Proportion of Variance Explained"
+    )
+  ) +
+
+  # Labels and title
+  labs(
+    title = "Scree Plot For Principal Components",
+    x = "Principal Component"
+  ) +
+  theme_minimal() +
+  theme(
+    plot.title = element_text(hjust = 0.5, face = "bold"),
+    plot.subtitle = element_text(hjust = 0.5, face = "italic"),
+    legend.position = "bottom",
+    legend.title = element_text(size = 12),
+    legend.text = element_text(size = 10),
+    axis.text = element_text(size = 10),
+    axis.title = element_text(size = 12, face = "bold")
+  )
+
+ggsave("../plots/scree_plot.png", width = 12, height = 8)
+```
+
+## 5.2 ROC Curves
+
+```{r}
+ggroc(list(
+  benchmark = roc_benchmark,
+  scree = roc_scree,
+  greaterthan1 = roc_greaterthan1
+)) +
+  labs(
+    title = "ROC Curves of Different Models",
+    x = "1 - Specificity (False Positive Rate)",
+    y = "Sensitivity (True Positive Rate)",
+    color = "Model"
+  ) +
+  geom_abline(
+    intercept = 1,
+    slope = 1,
+    linetype = "dashed",
+    color = "gray",
+    size = 0.8
+  ) +
+  scale_color_manual(
+    values = c("steelblue", "orange", "darkgreen"),
+    labels = c("Benchmark", "Scree", "Greater Than 1")
+  ) +
+  theme_minimal() +
+  theme(
+    plot.title = element_text(hjust = 0.5, face = "bold"),
+    plot.subtitle = element_text(hjust = 0.5, face = "italic"),
+    legend.position = "bottom",
+    legend.title = element_text(size = 12),
+    legend.text = element_text(size = 10),
+    axis.text = element_text(size = 10),
+    axis.title = element_text(size = 12, face = "bold")
+  )
+
+ggsave("../plots/roc_curves.png", width = 12, height = 8)
+```
+
+# 6. Iteratively Add Components
+
+## 6.1 Estimate Logistic Regression Models
+
+```{r}
+logit_results <- rep(NA, length(pc_greaterthan1)) # to store results
+logit_warnings <- rep(FALSE, length(pc_greaterthan1)) # to track warnings
+
+for (i in seq_along(pc_greaterthan1)) {
+  captured_warning <- NULL # set caputured_warning to NULL
+  logit <- suppressWarnings(
+    withCallingHandlers({
+      # Fit logistic regression model
+      logit <- data_pca %>%
+        select(all_of(pc_greaterthan1[1:i]), class) %>%
         train(
@@
           family = "binomial",
-    trControl = trainControl(method = "cv", number = 10)
+          trControl = trainControl(method = "LOOCV", savePredictions = "final", classProbs = TRUE)
+        )
+      # Store accuracy
+      logit_results[i] <- logit$results$Accuracy
+    }, warning = function(w) {
+      captured_warning <<- w # Capture the warning
+    })
   )
 
-summary(logit_7)
-logit_7$results
+  # logit_warnings == T if warning occurred
+  if (!is.null(captured_warning)) {
+    logit_warnings[i] <- TRUE
+    message(paste("Warning in iteration", i, ":", conditionMessage(captured_warning)))
+  }
+}
 ```
+
+## 6.2 Plot Accuracy by Number of Components
+
+```{r}
+# Create a data frame to combine all data
+plot_data <- data.frame(
+  x = 1:length(pc_greaterthan1),
+  y = logit_results,
+  warning = ifelse(logit_warnings, "No", "Yes") # Create a factor for warnings
+)
+
+ggplot() +
+  # Line for accuracy results
+  geom_line(data = plot_data, aes(x = x, y = y)) +
+
+  # Points with color mapped to 'warning'
+  geom_point(
+    data = plot_data,
+    aes(x = x, y = y, color = warning), size = 3
+  ) +
+
+  labs(
+    x = "Number of Components",
+    y = "Accuracy",
+    title = "Accuracy by Number of Principal Components",
+    color = "Convergence" # Label for the legend
+  ) +
+  scale_color_manual(values = c("No" = "darkblue", "Yes" = "darkorange")) + # Custom colors
+  theme_minimal() +
+  theme(
+    plot.title = element_text(hjust = 0.5, face = "bold"),
+    plot.subtitle = element_text(hjust = 0.5, face = "italic"),
+    legend.position = "bottom",
+    legend.title = element_text(size = 12),
+    legend.text = element_text(size = 10),
+    axis.text = element_text(size = 10),
+    axis.title = element_text(size = 12, face = "bold")
+  )
+
+ggsave("../plots/accuracy_by_components.png", width = 12, height = 8)
+```

### 2025-01-17 response_letter/response_letter.Rmd
Add all folders

---
title: 'Group assignment: response to peer reviews'
author: "Jonathan Koop, Martijn van Dam, Merlin Urbanski"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Dear Reviewers,

Thank you for your detailed and thoughtful feedback. Below is a brief summary of your main points and an overview of how we have addressed them in our revised report:

You recommended that we refine the structure of our report by shortening and professionalizing section titles and introducing subsections in the Methods section to improve clarity. We have implemented these suggestions, made the titles more concise and adding distinct subheadings, which now makes the report more coherent and better to read.

In the Introduction, you requested additional details on the dataset itself—such as whether the variables are numerical or categorical, their ranges, and what constitutes “high” or “low” values. We have expanded this section now, and provide examples and descriptive statistics to give readers a better sense of how the data are organized and what types of values are considered high or low.

You also suggested elaborating on the methods used by Cilia et al. (2022), particularly regarding the classifiers they employed. Although we initially considered comparing our PCA approach to theirs, we opted not to do so in order state a research problem within our word limit and the scope of the course. Instead, we have highlighted the rationale for using PCA as a dimensionality-reduction technique and clarified our benchmark approach, backed by relevant literature references.

Lastly, you emphasized the importance of evaluating sensitivity and specificity in the context of diagnosing Alzheimer’s disease. In response, we have now included an ROC curve in our report, which offers a more comprehensive view of these metrics and ensures a balanced assessment of our model’s diagnostic performance.

We appreciate your constructive input and thank you for your effort in reviewing our report.

Sincerely,

Jonathan, Martijn and Merlin

### 2025-01-17 code/analysis.qmd
Add all folders

@@
 predictions_model_benchmark <- model_benchmark$pred
-roc_benchmark <- roc(response = predictions_model_benchmark$obs, predictor = predictions_model_benchmark$P)
+(roc_benchmark <- roc(response = predictions_model_benchmark$obs, predictor = predictions_model_benchmark$P))
 ```
@@
 predictions_model_scree <- model_scree$pred
-roc_scree<- roc(response = predictions_model_scree$obs, predictor = predictions_model_scree$P)
+(roc_scree<- roc(response = predictions_model_scree$obs, predictor = predictions_model_scree$P))
 ```
@@
 predictions_model_greaterthan1 <- model_greaterthan1$pred
-roc_greaterthan1 <- roc(response = predictions_model_greaterthan1$obs, predictor = predictions_model_greaterthan1$P)
+(roc_greaterthan1 <- roc(response = predictions_model_greaterthan1$obs, predictor = predictions_model_greaterthan1$P))
 ```
